---
title: "End-to-End Omics Analysis & Capstone Report"
author: "Bioinformatics Core"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    theme: flatly
    code-fold: show
    code-summary: "Click to toggle code"
    self-contained: true
params:
  project_title: "Bulk RNA-Seq & Functional Enrichment Pipeline"
  padj_cutoff: 0.05
  lfc_cutoff: 1.0
---

```{r setup, include=FALSE}
# Chunk setup options
knitr::opts_chunk.param = list(
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 5,
  dpi = 300
)

library(dplyr)
library(ggplot2)
library(DT)
```

# Executive Summary

This report documents the end-to-end pipeline run for **`r params$project_title`**.

* **FDR Significance Threshold ($p_{adj}$):** `r params$padj_cutoff`
* **Log2 Fold-Change Threshold:** `r params$lfc_cutoff`

---

# Data Ingestion & Processing

```{r data-simulation}
set.seed(42)

# Generate synthetic differential expression results table
results_df <- data.frame(
  GeneSymbol = paste0("GENE_", 1:1000),
  baseMean   = runif(1000, 10, 5000),
  log2FC     = rnorm(1000, mean = 0, sd = 1.5),
  pvalue     = runif(1000, 0, 0.05)
) %>%
  mutate(
    padj = p.adjust(pvalue, method = "BH"),
    Significance = case_when(
      padj < params$padj_cutoff & log2FC > params$lfc_cutoff ~ "Up-regulated",
      padj < params$padj_cutoff & log2FC < -params$lfc_cutoff ~ "Down-regulated",
      TRUE ~ "Not Significant"
    )
  )
```

---

# Differential Expression Results

::: {.panel-tabset}

## Volcano Plot

```{r volcano-plot, fig.cap="Volcano plot displaying differential gene expression significance."}
ggplot(results_df, aes(x = log2FC, y = -log10(padj), color = Significance)) +
  geom_point(alpha = 0.7, size = 1.8) +
  scale_color_manual(values = c("Up-regulated" = "#D95F02", 
                                "Down-regulated" = "#7570B3", 
                                "Not Significant" = "grey70")) +
  geom_vline(xintercept = c(-params$lfc_cutoff, params$lfc_cutoff), linetype = "dashed") +
  geom_hline(yintercept = -log10(params$padj_cutoff), linetype = "dashed") +
  theme_minimal() +
  labs(
    title = "Differential Expression Volcano Plot",
    x = expression("log"[2]*" Fold Change"),
    y = expression("-log"[10]*" Adjusted p-value")
  )
```

## Significant DEGs Table

```{r deg-table}
sig_degs <- results_df %>% 
  filter(Significance != "Not Significant") %>%
  arrange(padj)

datatable(
  sig_degs,
  options = list(pageLength = 5, autoWidth = TRUE),
  caption = "Table of Significantly Differential Genes"
)
```

:::

# Conclusion & Next Steps

Pipeline execution completed successfully with **`r nrow(sig_degs)`** differentially expressed genes identified.